---
title: "r_ggplot2"
author: "Jian"
date: "24/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exercise 1
Install and Load the ggplot2 package
```{r}
renv::install('ggplot2')
library(ggplot2)
```

Run ggplot(diamonds). Returns a grey box because no axis is specified
```{r}
ggplot(diamonds)
```

Specify the axies. Return a grid plot because no 'geom' is specified
```{r}
ggplot(diamonds, aes(x=carat, y=price))
```

Draw points showing carat on the x-axis and price on the y-axis
```{r}
ggplot(diamonds,aes(x=carat,y=price))+
  geom_point()
```

Color data points by cut
```{r}
ggplot(diamonds,aes(x=carat,y=price,color=cut))+
  geom_point()
```
This following code is doing the same thing as above but the color is only applied to the geom_point() function
```{r}
ggplot(diamonds,aes(x=carat,y=price))+
  geom_point(aes(color=cut))
```


Add a smoothed mean trend line
```{r}
ggplot(diamonds,aes(x=carat,y=price))+
  geom_point(aes(color=cut))+
  geom_smooth(method='lm')
```


Exercise 2
Plot a histogram of chick weight
```{r}
plot1_histogram <- ggplot(ChickWeight,aes(x=weight))+
  geom_histogram(aes(fill=Diet),color='black',bins=50)+
  labs(title='histogram of chick weight',
       x='chick weight',
       y='number of chicks',
       subtitle = 'chick weight dataset by Diet')+
  theme(axis.title = element_text(size=16,face='bold'),
        axis.text.x = element_text(size = 10,angle = 30,hjust = 0,color = 'red'))+
  #ylim(NA,25)+ # NA is to give ggplot to decide the lower limit
  facet_wrap(~Diet,scales='free_y',ncol = 1)
print(plot1_histogram)
```

Exercise 3
Plot a scatter plot of chick weight over time
```{r}
plot2_scatter <- ggplot(ChickWeight,aes(x=Time,y=weight,color=Diet))+
  geom_point()+
  geom_smooth(method = 'lm')+
  labs(title = 'Scatter plot of chick weight over time',y='chick weight')+
  theme(axis.title = element_text(size = 15,face = 'bold'),
        axis.text = element_text(size=10))+
  scale_color_manual(values = c('1'='blue','2'='red','4'='orange','3'='green')) # Use scale_color_manual() to manually create the color
print(plot2_scatter)
```

Use gridExtra to concatenate
```{r}
renv::install('gridExtra')
library(gridExtra)
```

```{r}
plot1_histogram_2 <- arrangeGrob(plot1_histogram,top = 'A') # Name plot1_histogram as 'A' and assign to plot1_histogram_2
plot2_scatter_2 <- arrangeGrob(plot2_scatter,top = 'B') # Name plot2_scatter as 'B' and assign to plot1_scatter_2
grid.arrange(plot1_histogram_2,plot2_scatter_2,ncol=2,nrow=1)
```

Use cowplot to concatenate
```{r}
renv::install('cowplot')
library(cowplot)
```


```{r}
plot_grid(plot1_histogram,plot2_scatter,labels = c('A','B'), ncol=2, nrow=1)
```

Exercise 4
Use RColorBrewer to manually specify the color
```{r}
ggplot(ChickWeight,aes(x=weight))+
  geom_histogram(aes(fill=Diet),color='black',bins=50)+
  scale_fill_brewer(palette = 'RdYlGn')+ # This is the use the color code from the RColorBrwer
  labs(title='histogram of chick weight',
       x='chick weight',
       y='number of chicks',
       subtitle = 'chick weight dataset by Diet')+
  theme(axis.title = element_text(size=16,face='bold'),
        axis.text.x = element_text(size = 10,angle = 30,hjust = 0,color = 'red'))+
  #ylim(NA,25)+ # NA is to give ggplot to decide the lower limit
  facet_wrap(~Diet,scales='free_y',ncol = 1)
```

